Evolutionary Design of Particles for Collectively Intelligent Problem Solving

### 2014-05-08 docs/paper.tex
results

\documentclass[12pt,journal,compsoc]{IEEEtran}
\providecommand{\PSforPDF}[1]{#1}
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage{graphicx}
\usepackage{float}
\floatstyle{boxed}
\restylefloat{figure}

@@

@@

\begin{abstract}

\end{abstract}

@@

\section{Results}
results

\section{Discussion}
This work described herein represents only the first steps in the application of evolutionary techniques to this problem domain. By mutating and applying selective pressure to the movement component parameters, we were able to show that a machine could optimize the parameters better than a person. This was a form of optimization, but it did not result in novel behaviors. Indeed, since we hard-wired the transitions between the states (allowing for the guarding threshold to evolve) and allowed the evolver to mutate existing movement components in a given state but not to add them (i.e., we pre-defined which components could have non-zero weight for each state), we held constant the "meaning" of each state.

>>>>

\documentclass[12pt,journal,compsoc]{IEEEtran}
\providecommand{\PSforPDF}[1]{#1}
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{graphicx}
\usepackage{float}
\floatstyle{boxed}
\restylefloat{figure}

@@

\IEEEcompsoctitleabstractindextext{%
\begin{abstract}

\end{abstract}}

@@

\IEEEdisplaynotcompsoctitleabstractindextext

@@

\section{Results}

In this section we will evaluate the effectiveness of both evolved flocking particles vs. human created particles, as well as the ability of an evolutionary computation to design such particles. In short, the evolutionary computation mechanism was able to design a particle system that exhibited variations in behaviors that were different enough from the human designed particles so as not only be competitive with the human design team, but able to defeat the other team in terms of resources collected.

Table \ref{table:meanfitness} shows the mean, standard deviation, maximum, and minimum resources collected for the black team over ten runs of the simulation with the same parameters. Black teams of comparsion are the human designed team, the best evolved team, and a random team. The evolution designed team performs significantly better than the human designed team.

\begin{table}
    \renewcommand{\arraystretch}{1.5}
    \centering
    \begin{tabular}{l | c c cc}
        \hline
        Simulation & Mean & Std Dev & Max & Min \\
        \hline
        human & 119.0 & 43.2 & 274 & 119 \\
        evolved & 296.0 & 25.6 & 336 & 254 \\
        random & 0.3 & 0.67 & 2 & 0 \\
        \hline
    \end{tabular}
    \caption{Aggregated fitness of various particles competing with human designed particles.}
    \label{table:meanfitness}
\end{table}

The average fitness over time of the black and red teams is displayed in \ref{fig:head2head} which graphs the mean fitness of ten individual runs of the best human designed team (the red team) vs the best evolved team, the black team. Within 6000 timesteps the red team begins to lose resources as the black team exploits the red base in addition to more resource locations than the red team, which favors a strategy of exploiting one resource deposit at a time.

\begin{figure}[h!]
    \centering
        \includegraphics[width=0.5\textwidth]{figures/head2head}
    \caption{Performance of evolved particles vs human designed particles in head to head simulation of 10,000 timesteps}
    \label{fig:head2head}
\end{figure}

The performance of our two evolutionary computations is expressed in \ref{fig:meanfit1} and \ref{fig:meanfit2} which plots the mean fitness of the entire population per each generation. In experiment one, only the mutation operator was used to generate the next population. In experiment two, both recombination and mutation were used to produce the next generation.

\begin{figure}[h!]
    \centering
        \includegraphics[width=0.5\textwidth]{figures/meanfit_experiment1}
    \caption{Mean fitness per generation with mutation}
    \label{fig:meanfit1}
\end{figure}

\begin{figure}[h!]
    \centering
        \includegraphics[width=0.5\textwidth]{figures/fitness21}
    \caption{Mean fitness per generation with mutation and recombination}
    \label{fig:meanfit2}
\end{figure}

As is shown from the graphs; the first experiment was wildly unstable in terms of its search across the optimal state space. This may have been due to the fact that highly variable mutation caused dramatic shifts in the behavior of the phenotypes. It also may have been due to a programming error on the part of the research team. In the second evolution with recombination as well as mutation, there is more steadily growing fitness over time and a more stable population to select from to apply genetic operations.

\section{Discussion}
The evolution designed particles surpassed their human design counterparts through a qualitatively novel mechanism; the human designed team prefered to stay in one or two larger flocks; whereas the high performing evolved teams split into more, smaller flocks and were able to exploit the search space more effectively.

This work described herein represents only the first steps in the application of evolutionary techniques to this problem domain. By mutating and applying selective pressure to the movement component parameters, we were able to show that a machine could optimize the parameters better than a person. This was a form of optimization, but it did not result in novel behaviors. Indeed, since we hard-wired the transitions between the states (allowing for the guarding threshold to evolve) and allowed the evolver to mutate existing movement components in a given state but not to add them (i.e., we pre-defined which components could have non-zero weight for each state), we held constant the "meaning" of each state.
